Effectiveness of BNT162b2 booster doses in England: an observational study in OpenSAFELY-TPP

### 2022-03-28 manuscript/draft-manuscript.Rmd
Update draft-manuscript.Rmd

We matched `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls. During the first 28 days after the booster, estimated booster effectiveness comparing a BNT162b2 booster dose to two doses only was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display95)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.

@@

In mid-September 2021, the national COVID-19 vaccination programme in England administered its first booster doses in adults who had already received their two-dose primary vaccination course @NHSEnglandNHS. Eligibility was initially restricted to those at highest risk of severe disease, then progressively extended. By 15 December 2021 every adult was eligible @NHSEnglandNHSb. Booster doses were initially available no earlier than six months after dose two, but this was reduced to three months on 8 December 2021, following concerns over the emergence of the Omicron variant @NHSEnglandNHSa1 @NHSEnglandNHSa2. Vaccine prioritisation schedules were guided by recommendations from the Joint Committee for Vaccine and Immunisation (JCVI) expert working group.

@@

All data were linked, stored and analysed securely within the OpenSAFELY platform: <https://opensafely.org/>. Primary care records managed by the GP software provider TPP were linked, using NHS numbers, to A&E attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS). COVID-19 vaccination history is available in the GP record directly via the National Immunisation Management System (NIMS). Data include pseudonymized data such as coded diagnoses, medications and physiological parameters. No free text data are included.

@@

We included adults aged $\geq$ 18 years who were eligible for booster vaccination between `r format(as.Date(study_dates$pfizerstart_date), "%e %B")` and `r format(as.Date(study_dates$pfizerend_date), "%e %B %Y")` inclusive (the "recruitment period"). On each day within this period, eligibility for the study was determined as follows: registered at a GP practice using TPP's SystmOne clinical information system; received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and Moderna mRNA-1273 were not considered due to small numbers); not a health or social care worker, not resident in a care or nursing home; not medically housebound or receiving end-of-life care; no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; not undergoing an unplanned hospital admission; complete information on sex, ethnicity, deprivation, and NHS region.

Additionally, if the rolling weekly average count of BNT162b2 booster doses within strata defined by region, JCVI group, and the week of second dose was below 50 then recruitment did not occur for that stratum on that day. This helped ensure that recruitment was restricted to those boosted in line with national prioritisation schedules and that the matched controls were both eligible and able to receive a booster dose at the time of potential recruitment.

@@

The supplementary materials provide further details of the matching process.

@@

The supplementary materials provide further details of the codelists and data sources used for all variables in the study.

@@

We estimated Kaplan-Meier cumulative incidence curves separately in the boosted and unboosted groups, and the risk difference and risk ratio at `r label_number(accuracy=1)(max(postbaselinecuts)/7)` weeks.

We used Cox models, stratified by recruitment day, NHS region, and brand of second vaccine dose, to estimate hazard ratios (HRs) comparing boosted with unboosted people, overall and separately for days 1-28 and 29-70, and also for days 1-7, 8-14, 15-28, 29-42, and 43-70. As we were not able to match on all potential confounders, these models included the following additional covariates: sex (male or female); English Index of Multiple Deprivation (IMD, grouped by quintile); ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census); prior morbidity count (diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease); learning disabilities; severe mental illness; the number of SARS-CoV-2 tests in the 6 months prior to the study start date; elective in-hospital episode at the time of recruitment.A robust variance estimator was used to account for inclusion of some people in both the boosted and unboosted groups. Booster vaccine effectiveness was calculated as one minus the hazard ratio, expressed as a percentage.

@@

We also estimated booster effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or $\geq$ 65 years); clinically extremely vulnerable or not.

## Missing data

After exclusions for missing values on demographic variables, there were no missing values in the remaining variables as they were each defined by the presence or absence of clinical codes or events in the patient record.

@@

There were `r label_number(1, big.mark=",")(filter(km70, treatment=="pfizer", outcome=="postest") %>% pull(cml.events))` positive SARS-CoV-2 tests,
`r label_number(1, big.mark=",")(filter(km70, treatment=="pfizer", outcome=="covidadmitted") %>% pull(cml.events))` COVID-19 hospital admissions, 
`r label_number(1, big.mark=",")(filter(km70, treatment=="pfizer", outcome=="coviddeath") %>% pull(cml.event))` COVID-19 deaths, 
and `r label_number(1, big.mark=",")(filter(km70, treatment=="pfizer", outcome=="noncoviddeath") %>% pull(cml.events))` non-COVID-19 deaths 
across `r label_number(1, big.mark=",")((filter(km70, treatment=="pfizer", outcome="coviddeath") %>% pull(cml.atrisk)) /7)` person-weeks of follow-up. 
The 10-week risk of positive SARS-CoV-2 test was `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` per 1000 in the boosted group 
and `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` in the unboosted group, 
representing a risk difference of `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=1000))` per 1000. 
For COVID-19 hospital admission, this was `r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` 
and `r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` 
respectively (`r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=-1000))` fewer events in the boosted group). 
For COVID-19 death it was `r km70 %>% filter(treatment=="pfizer", outcome=="coviddeath", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` 
and `r km70 %>% filter(treatment=="pfizer", outcome=="coviddeath", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` 
(`r km70 %>% filter(treatment=="pfizer", outcome=="coviddeath", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=-1000))` fewer events), 
and for non-COVID-19 death it was `r km70 %>% filter(treatment=="pfizer", outcome=="noncoviddeath", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` 
and `r km70 %>% filter(treatment=="pfizer", outcome=="noncoviddeath", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` 
(`r km70 %>% filter(treatment=="pfizer", outcome=="noncoviddeath", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=-1000))` fewer events) respectively  (Table 2).

@@

The overall (days 1-70) estimated booster effectiveness comparing the boosted and unboosted groups was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display95)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death. Estimated booster effectiveness against COVID-19 outcomes was generally lower during days 1-28 than days 29-70: `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for positive SARS-CoV-2 test; `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` respectively for COVID-19 hospital admission; `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` respectively for COVID-19 death. Estimated booster effectiveness against non-COVID19 death was similar during days 1-28 and 29-70 (`r filter(meta2hazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)`).

Estimated effectiveness for narrower time periods are provided in supplementary materials (Table S3).

@@

Amongst people aged $\geq$ 65 years estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.  Corresponding effectiveness was slightly lower in people aged 18-64 years: `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` respectively.

@@

#### Table 3: Estimated booster vaccine effectiveness

Effectiveness $(100 \times (1-\text{hazard ratio}))$ for the main and subgroup analyses and for all outcomes.

@@

#### Figure 2: Cumulative outcome incidence

Kaplan-Meier estimates of cumulative incidence in matched boosted and unboosted treatment groups, stratified by primary course and without further adjustment for potential confounders.

@@

#### Figure 3: Estimated booster effectiveness

For each outcome based on the fully adjusted model, the hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course and time since boosting. Models with less extensive confounder adjustment are provided in supplementary materials (Figure S3).

>>>>

We matched `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls. Estimated booster effectiveness comparing a BNT162b2 booster dose to two doses only was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display95)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death. Estimated effectiveness was similar in those receiving BNT162b2 or ChAdOx1-S two-dose primary vaccination course, but slightly lower in those classified as clinically extremely vulnerable (`r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death) and in those aged 18-65 years (`r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` respectively). Period-specific effects were also estimated, showing some evidence of waning booster effectiveness against positive SARS-CoV-2 tests (`r filter(hazards, outcome=="postest", treatment=="pfizer", term_right=28, subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` at 15-28 days versus `r filter(hazards, outcome=="postest", treatment=="pfizer", term_right=70, subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` at 43-70 days).

@@

In mid-September 2021, the national COVID-19 vaccination programme in England administered its first booster doses in adults who had already received their two-dose primary vaccination course @NHSEnglandNHS. Eligibility was initially restricted to those at highest risk of severe disease, then progressively extended. By 15 December 2021 every adult was eligible @NHSEnglandNHSb. Booster doses were initially available no earlier than six months after dose two, but this was reduced to three months on 8 December 2021, following concerns over the emergence of the Omicron variant @NHSEnglandNHSa1 @NHSEnglandNHSa2. Vaccine prioritisation schedules were guided by recommendations from the Joint Committee for Vaccine and Immunisation (JCVI) expert working group @covid19.

@@

All data were linked, stored and analysed securely within the OpenSAFELY platform: <https://opensafely.org/>. Primary care records managed by the GP software provider TPP were linked, using NHS numbers, to A&E attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS). COVID-19 vaccination history is available in the GP record directly via the National Immunisation Management System (NIMS).

@@

We included adults aged \ge 18 years who were eligible for booster vaccination between `r format(as.Date(study_dates$pfizerstart_date), "%e %B")` and `r format(as.Date(study_dates$pfizerend_date), "%e %B %Y")` inclusive (the "recruitment period"). On each day within this period, eligibility for the study was determined as follows: registered at a GP practice using TPP's SystmOne clinical information system; received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and Moderna mRNA-1273 were not considered due to small numbers); not a health or social care worker, not resident in a care or nursing home; not medically housebound or receiving end-of-life care; no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; not undergoing an unplanned hospital admission; complete information on sex, ethnicity, deprivation, and NHS region.

Additionally, if the rolling weekly average count of BNT162b2 booster doses within strata defined by region, JCVI priority group, and the week of second dose was below 50 then recruitment did not occur for that stratum on that day. This helped to ensure that recruitment was restricted to those eligible and able to receive a booster dose in line with national prioritisation schedules at the time of potential recruitment.

@@

@@

@@

We estimated Kaplan-Meier cumulative incidence curves separately in the boosted and unboosted groups, and the corresponding 10-week risk difference and risk ratio.

We used Cox models, stratified by recruitment day, NHS region, and two-dose primary vaccine course, to estimate hazard ratios (HRs) comparing boosted with unboosted people, overall and separately for days 1-28 and 29-70, and also for days 1-7, 8-14, 15-28, 29-42, and 43-70. As we were not able to match on all potential confounders, these models included the following additional covariates: sex (male or female); English Index of Multiple Deprivation (IMD, grouped by quintile); ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census); prior morbidity count (diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease); learning disabilities; severe mental illness; the number of SARS-CoV-2 tests in the 6 months prior to the study start date; elective in-hospital episode at the time of recruitment.A robust variance estimator was used to account for inclusion of some people in both the boosted and unboosted groups. Booster vaccine effectiveness was calculated as one minus the hazard ratio, expressed as a percentage.

@@

We also estimated booster effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or \ge 65 years); clinically extremely vulnerable or not; evidence of prior SARS-CoV-2 infection or not.

<!-- ## Missing data -->

<!-- After exclusions for missing values on demographic variables, there were no missing values in the remaining variables as they were each defined by the presence or absence of clinical codes or events in the patient record.  -->

@@

There were `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` positive SARS-CoV-2 tests,
`r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, 
`r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths, 
and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="noncovidadmitted", fup_period=="All") %>% pull(events))` non-COVID-19 deaths 
across `r label_number(1, big.mark=",")((filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(yearsatrisk)) * 365.25/7)` person-weeks of follow-up. 
The 10-week risk of positive SARS-CoV-2 test was `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` per 1000 in the boosted group 
and `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` in the unboosted group, 
representing a risk difference of `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=1000))` per 1000. 
For COVID-19 hospital admission, this was `r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` 
and `r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` 
respectively (`r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=-1000))` fewer events in the boosted group). 
For COVID-19 death it was `r km70 %>% filter(treatment=="pfizer", outcome=="coviddeath", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` 
and `r km70 %>% filter(treatment=="pfizer", outcome=="coviddeath", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` 
(`r km70 %>% filter(treatment=="pfizer", outcome=="coviddeath", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=-1000))` fewer events), 
and for non-COVID-19 death it was `r km70 %>% filter(treatment=="pfizer", outcome=="noncoviddeath", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` 
and `r km70 %>% filter(treatment=="pfizer", outcome=="noncoviddeath", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` 
(`r km70 %>% filter(treatment=="pfizer", outcome=="noncoviddeath", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=-1000))` fewer events) respectively  (Table 2).

@@

The overall (days 1-70) estimated booster effectiveness comparing the boosted and unboosted groups was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display95)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death (Table 3). Estimated booster effectiveness against COVID-19 outcomes was generally lower during days 1-28 than days 29-70: `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for positive SARS-CoV-2 test; `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` respectively for COVID-19 hospital admission; `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` respectively for COVID-19 death. Estimated booster effectiveness against non-COVID19 death was similar during days 1-28 and 29-70 (`r filter(meta2hazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)`).

@@

Amongst people aged \ge 65 years estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.  Corresponding effectiveness was slightly lower in people aged 18-64 years: `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` respectively.

@@

#### Table 3:  Estimated booster vaccine effectiveness ($100 \times (1-\text{hazard ratio})$) for the main and subgroup analyses and for all outcomes.

@@

#### Figure 2: Kaplan-Meier estimates of cumulative incidence in matched boosted and unboosted treatment groups, stratified by primary course and without further adjustment for potential confounders.

@@

#### Figure 3: Estimated booster vaccine effectiveness ($100 \times (1-\text{hazard ratio})$) for each outcome based on the fully adjusted model, stratified by primary course and time since boosting. Models with less extensive confounder adjustment are provided in supplementary materials (Figure S3).
